Learning and Reputation under Monopoly and Duopoly: \\A Competing Bandits Approach

### 2018-08-27 paper/competing_bandits.tex
fixes to the paper

We evaluate the performance of the different learning algorithms on the bandit priors we consider in isolation. There are two reasons why this is important to consider. First, we want to confirm that the reputation ordering of the algorithms is what we would expect according to the multi-armed bandit literature where $TS > DEG > DG$ for sufficiently large $t$. Second, we want to better understand what statistics of the instance we can look at in isolation in order to help us predict and understand what to expect in the competition game.

@@

\begin{finding}
\textit{The mean performance of an algorithm is not a sufficient statistic for understanding the performance of an algorithm in competition. Looking at the distribution of reputation difference between algorithms we find that it tend is skewed to the right. Additionally we show that the mean performance of the algorithms is not always predictive of the competition game, even when both algorithms are given a large number of agents for free.}
\end{finding}

@@

To better capture the comparison between two algorithms we explicitly plot the distribution of the reputation difference between $TS$ and $DG$. This figure seems to confirm the intuition noted previously since the reputation distribution has its largest mass around the point just below 0 but it is skewed to the right even as $t$ gets large. Given the skewedness of the distribution, the mean is not a representative statistic of the entire reputation difference distribution. There are alternative statistics that we could consider such as the median or numerically calculating $\Pr(reputation(TS) - reputation(DG)) > 0$ using the estimated density, but instead we define a simple and natural statistic to help us reason about what happens under competition, the relative reputation proportion.

@@

We define the \textit{relative reputation proportion} in order to better understand what happens under competition.

@@

\begin{definition}
\textit{Relative reputation cost and benefit} - the relative reputation loss an algorithm incurs from purposeful exploration compared to the greedy alternative. Here we treat it as an empirical definition based on the relative reputation proportion where the ``cost" regime is when the proportion $< 0.5$ and the ``benefit" regime is when the proportion is $> 0.5$ for the better algorithm. We call the priors where there is an early costly period followed by a later benefit period, relative reputation costly priors.
\end{definition}

@@

Table \ref{sim_unif} shows the results under the Uniform prior. For the low values of the warm start we can see that $DG > TS$, but that for larger warm starts, $TS > DG$ \footnote{The same results hold for the Heavy Tail prior and are reported in the supplemental material}. Looking at the relative reputation plots in Figure \ref{relative_rep_plots}, we can interpret fixing a warm start value as fixing the starting point on the relative reputation plots. The proportion of first rounds in the competition game that will go to a firm playing alg $A$ over a firm playing alg $B$ will correspond to the relative reputation proportion at time $k$. The intuition for the result in competition aligns with the fact that for Uniform and Heavy Tail, $TS$ is relative reputation costly compared to $DG$ and so, for low warm starts, $TS$ has not yet recovered from the relative reputation loss due to purposeful exploration and so $DG$ beats $TS$. However, if we move the warm start sufficiently high so that $TS$ moves to the relative reputation benefit regime, $TS > DG$.

However, we do not always see that $DG > TS$ under some warm starts. Table \ref{sim_nih} shows the results for the Needle In Haystack prior where we see that $TS > DG$. The intuition for this is precisely that, under the Needle In Haystack, $TS$ is not relative reputation costly compared to $DG$ and thus we see that $TS > DG$.

@@

\begin{finding}
\textit{When the incumbent plays $TS$, the data advantage serves as a larger barrier to entry compared to reputation. However, when the incumbent plays $DEG$ or $DG$ it does not} 
\end{finding}

>>>>

We evaluate the performance of the different learning algorithms on the bandit priors we consider in isolation. There are two reasons why this is important to consider. First, we want to confirm that the reputation ordering of the algorithms is what we would expect according to the multi-armed bandit literature where, according to the mean reputation, $TS > DEG > DG$ for sufficiently large $t$. Second, we want to better understand what statistics of the priors we can look at in isolation in order to help us predict and understand what to expect in the competition game.

@@

\begin{finding}
\textit{The mean performance of an algorithm is not a sufficient statistic for understanding the performance of an algorithm in competition. Looking at the distribution of reputation difference between algorithms we find that it tend is skewed to the right. Additionally we provide a counterexample where the mean performance does not predict the result of the competition game.}
\end{finding}

@@

To better capture the comparison between two algorithms we explicitly plot the distribution of the reputation difference between $TS$ and $DG$. This figure seems to confirm the intuition noted previously since the reputation distribution has its largest mass around the point just below 0 but it is skewed to the right even as $t$ gets large. Given the skewedness of the distribution, the mean is not a representative statistic of the entire reputation difference distribution. There are alternative statistics that we could consider such as the median or numerically calculating $\Pr(reputation(TS) - reputation(DG)) > 0$ using the estimated density, but instead we define a simple and natural statistic to help us reason about what happens under competition, the\textit{relative reputation proportion}.

@@

@@

\begin{definition}
\textit{Relative reputation cost and benefit} - the relative reputation loss an algorithm incurs from purposeful exploration compared to the greedy alternative. Here we treat it as an empirical definition based on the relative reputation proportion where the ``cost" regime is when the proportion $< 0.5$ and the ``benefit" regime is when the proportion is $> 0.5$ for the better algorithm. We call the priors where there is an early costly period followed by a later benefit period relative reputation costly priors.
\end{definition}

@@

Table \ref{sim_unif} shows the results under the Uniform prior. For the low values of the warm start we can see that $DG > TS$, but that for larger warm starts, $TS > DG$ \footnote{The same results hold for the Heavy Tail prior and are reported in the supplemental material}. Looking at the relative reputation plots in Figure \ref{relative_rep_plots}, we can interpret fixing a warm start $k$ as fixing the starting point on the relative reputation plots. The proportion of first rounds in the competition game that will go to a firm playing alg $A$ over a firm playing alg $B$ will correspond to the relative reputation proportion at time $k$. The intuition for the result in competition aligns with the fact that, for Uniform and Heavy Tail, $TS$ is relative reputation costly compared to $DG$ and so, for low warm starts, $TS$ has not yet recovered from the relative reputation loss due to purposeful exploration and so $DG$ beats $TS$. However, if we move the warm start sufficiently high so that $TS$ moves to the relative reputation benefit regime, $TS > DG$.

However, we do not always see that $DG > TS$ under some warm starts. Table \ref{sim_nih} shows the results for the Needle In Haystack prior where we see that $TS > DG$. This is precisely due to the fact that, under the Needle In Haystack, $TS$ is not relative reputation costly compared to $DG$ and thus we see that $TS > DG$.

@@

\begin{finding}
\textit{When the incumbent plays $TS$, the data advantage serves as a larger barrier to entry compared to reputation. However, when the incumbent plays $DEG$ or $DG$ there is no substantial difference between the two.} 
\end{finding}
